Seminar: Programming models and Code Generation\\ Winter Term 2014/2015\\ {\bf Quasi deterministic parallel programming using LVars}

### 2015-01-05 draft.tex
new

To keep track of the running callbacks, LVish includes an auxiliary form of the above pattern taking additional arguments: the set of running callbacks $B$ and a set of values $H$ in the event set for which callbacks have already been launched. It detects quiescence by checking: Every event of interest that has occurred must be handled (be in $H$). Second, all existing callback threads must have completed and terminated with a value. If yes, then LVars state is freezed. To ensure quiescence is a permanent state, rather than a transient one, freezing is usually the very last step of an algorithm, permitting its result to be extracted. The $runParThenFreeze$ function inside LVish library does this, and thereby guarantees full determinism.\\ \\
\textbf{\textit{Proof of Quasi determinism for Lvish:}} According to the paper\cite{lkuper2} the main theorem says that if two executions starting from a configuration $\sigma$ terminate in configurations $\sigma{'}$ and $\sigma{''}$, then either $\sigma{'}$ and $\sigma{''}$ are the same configuration (up to a permutation on locations), or one of them is \emph{error}.\\
\textit{Independence property:} Like LVars, the LVish model uses the frame property again to capture the idea that independent effects commute with each other but this time with some restrictions. If we know that a \emph{store} $S$ evaluating expression $e$ steps to a new \emph{store} $S{'}$ with $e{'}$ then it is valid to say that $S\sqcup_{S}S''$ evaluating $e$ steps to a new store $S'\sqcup_{S}S''$ with $e'$ if the following conditions are satisfied:
\begin{itemize}
\item The status bits of $S'\sqcup_{S}S''$ and S agree with each other for all locations shared between them.
\item Locations in $S''$ cannot share the same name with locations newly allocated during the transition from one configuration
to another.
\end{itemize}
Please refer to the original paper and the technical report for elaborate details on lemmas and their proofs.

@@

element is $\bot$. Lattices for which every element is the lub of some set of atoms are called atomistic, and most application-specific lattices used by LVish programs have this property and the corresponding data structure usually exposes operations that
work at the atom level, semantically limiting puts to atoms, gets to threshold sets of atoms, and event sets to sets of atoms. A compact way to represent a change to the lattice is deltas allowing to easily and efficiently communicate such changes between
$puts$ and $gets$/handlers. For a set, a delta is an element; for a map, a key/value pair.\par
\textit{Idempotence,} meaning that $d\sqcup{d=d}$ for any element $d$. Repeated puts or freezes have no effect, if the scheduler is allowed to occasionally duplicate work, it is possible to save on synchronization costs. Since LVish computations are guaranteed to be idempotent, we could use such a scheduler (standard Chase-Lev deque\cite{sched}). Idempotence helps to deal with races between $put$ and $get/addHandler$. \\ \\
Internally $par$ monad represents computations in continuation passing style, in terms of their interpretation in the IO monad. The status bit of an LVar is tied together with a bag of waiting listeners, which include blocked gets and handlers.

@@

flow of a program or precisely the flow of values to expressions. The figure on the following page shows performance data[7] for LVish on 2 benchmarks. One is the blur benchmark based on k=2. The implementation using LVish for the 2 benchmarks uses lock-free data structures. Another synthetic benchmark is notChain which is simply a chain of 300 not functions that negates the benifit of the sharing approach in the LVish model.\\ \\

@@

many of the single assignment models, thereby representing a larger class of problems which can be effciently parallelized. The performance with regards to certain benchmark applications are an evidence of the same. \cite{newton1}discusses the viability of Lvish for certain graph algorithms, its drawbacks and proposed extension. Some of the drawbacks to highlight are the fact that Lvish side  effects are only monotonic updates and if a data structure needs to be updated multiple times it will have to be done out of place thereby frequently allocating LVars becoming a source of additional memory allocation and ineffcient relative to in-place versions. Blocking $get$ semantics could slow down until the dependency is met reducing the performance, hence instead
of blocking, the tasks could be retried at a later time/rescheduled when there dependency has been satisfied. The overhead in going through the scheduler many times has also been highlighted in[4] for applications like Maximum independent set. The
paper proposes a solution similar to an abort/retry feature.

>>>>

To keep track of the running callbacks, LVish includes an auxiliary form of the above pattern taking additional arguments: the set of running callbacks $B$ and a set of values $H$ in the event set for which callbacks have already been launched. It detects quiescence by checking: Every event of interest that has occurred must be handled (be in $H$). Second, all existing callback threads must have completed and terminated with a value. If yes, then LVars state is freezed. To ensure quiescence is a permanent state, rather than a transient one, freezing is usually the very last step of an algorithm, permitting its result to be extracted. The $runParThenFreeze$ function inside LVish library does this, and thereby guarantees full determinism.\\ \\
\textbf{\textit{Proof of Quasi determinism for Lvish:}} According to the paper\cite{lkuper2} the main theorem says that if two executions starting from a configuration $\sigma$ terminate in configurations $\sigma{'}$ and $\sigma{''}$, then either $\sigma{'}$ and $\sigma{''}$ are the same configuration (up to a permutation on locations), or one of them is \emph{error}.\\
\textit{Independence property:} Like LVars, the LVish model uses the frame property again to capture the idea that independent effects commute with each other but this time with some restrictions. If we know that a \emph{store} $S$ evaluating expression $e$ steps to a new \emph{store} $S{'}$ with $e{'}$ then it is valid to say that $S\sqcup_{S}S''$ evaluating $e$ steps to a new store $S'\sqcup_{S}S''$ with $e'$ if the following conditions are satisfied:
\begin{itemize}
\item The status bits of $S'\sqcup_{S}S''$ and S agree with each other for all locations shared between them.
\item Locations in $S''$ cannot share the same name with locations newly allocated during the transition from one configuration
to another.
\end{itemize}
For details on lemmas and their proofs, refer \cite{kuper}\cite{lkuper2}.

@@

element is $\bot$. Lattices in which every element is the LUB of some set of atoms are called atomistic, and most application-specific lattices used by LVish programs have this property. \ref{Figure1} is an example for an atomistic lattice where the atoms are singleton elements and all the elements can be obtained as LUB of these atoms. Data structures corresponding to such type of lattices usually expose operations that work at the atom level, semantically limiting $puts$ to atoms, $gets$ to threshold sets of atoms, and event sets to sets of atoms. LVish facilitates good performance for atomistic lattices by associating LVars with a set of deltas(changes), as well a lattice. For atomistic lattices, the deltas are just the atoms, for a set, a delta is an element; for a map, a key/value pair. Deltas allow easy and efficient communication of changes between $puts$ and $gets$/handlers. \par
\textit{Idempotence,} meaning that $d\sqcup{d=d}$ for any element $d$. Repeated $puts$ or $freezes$ have no effect, if the scheduler is allowed to occasionally duplicate work, it is possible to save on synchronization costs. LVish computations are guaranteed to be idempotent and uses the scheduler(\textit{standard Chase-Lev deque}). Idempotence helps to deal with races between $put$ and $get/addHandler$. \\ 
The status bit of an LVar is tied together with a bag of waiting listeners, which include blocked $gets$ and handlers.

@@

flow of a program or precisely the flow of values to expressions. \ref{Figure2} shows performance data for LVish implementation using lock-free data structures on 2 benchmarks. One is the blur benchmark based on k=2. Another is the synthetic benchmark notChain which is simply a chain of 300 not functions that negates the benefit of the sharing approach in the LVish model.\\ \\

@@

many of the single assignment models, thereby representing a larger class of problems which can be effciently parallelized. The performance with regards to certain benchmark applications are an evidence of the same. \cite{newton1}discusses the viability of Lvish for certain graph algorithms, its drawbacks and proposed extension. It says that Lvish side effects are only monotonic updates and if a data structure needs to be updated multiple times it will have to be done out of place thereby frequently allocating LVars resulting in a source of additional memory allocation and inefficient relative to in-place versions. Blocking $get$ semantics could slow down the performance until the dependency is met. So instead of blocking, the tasks could be retried at a later time/rescheduled when there dependency has been satisfied. The overhead in going through the scheduler many times has also been highlighted in\cite{newton1} for applications like Maximum independent set and also proposes a solution similar to an abort/retry feature.
